Simulating an Isolated Dwarf Galaxy with Energetic Feedback and Chemical Yields from Individual Stars

### 2018-03-17 main.tex
final revisions

Self-shielding of gas against the metagalactic UV background is important in high-resolution simulations, particularly for low-mass, low-metallicity dwarf galaxies where the UV background is capable of gradually photoevaporating unshielded gas from the galaxy \citep{Simpson2013}. We have implemented the \citet{Rahmati2013} approximate self-shielding method in \texttt{Grackle} to account for \ion{H}{1} self-shielding against the UV background \citep[see][ for more details of this implementation]{GrackleMethod}. We assume \ion{He}{1} ionization generally follows \ion{H}{1}. This allows us to approximate \ion{He}{1} self-shielding using the same form (A. Rahmati, private communication). We ignore \ion{He}{2} photoionization from the UV background entirely. Accounting for self-shielding in this manner leads to an inconsistency in using tabulated, optically-thin metal line cooling rates from Cloudy (see Section 4.1.1 of \citet{Hu2017}). As mentioned previously, we have re-computed metal line cooling tables using \texttt{Cloudy} models of optically thick clouds to be consistent with our self-shielding prescription. This is described in more detail in Appendix ~\ref{appendix:cooling}.

@@

To aid in comparison to other works, we define our ISM phases here, as adopted and modified from \citet{Draine2011} Table 1.3 and used consistently throughout this analysis. By construction, these phases are mutually exclusive. In the below, $f_{\rm H_2}$ is the molecular hydrogen fraction of the total gas mass.

@@

\section{Stellar Radiation Properties}
\label{appendix:radiation}
We detail the radiation properties of our model stars as a function of stellar mass and metallicity. The stellar radiation properties are determined from the OSTAR2002 grid \citep{Lanz2003} in the regime that it covers, or integrated from an adjusted black body curve for other stars, given the ZAMS stellar radius obtained from the PARSEC \citep{Bressan2012,Tang2014} stellar evolution data set. In Figure~\ref{fig:stellar radiation properties} we plot these properties. We use a constant factor across metallicities for stars not sampled on the OSTAR2002 grid to shift the black body radiation fluxes to be roughly continuous with the ionizing photon rates and luminosities as a function of stellar mass. This requires two factors for each radiation type, one for low mass and one for high mass stars. We use the following multiplicative factors to adjust the black body spectrum for \ion{H}{1} and \ion{He}{1} radiation respectively: [0.1, 3.2] and [0.0001, 4.0]. We do not find this adjustment to be necessary for the FUV and Lyman-Werner radiation bands.

The ionizing radiation photon energies are taken as the average ionizing photon energy for a black body of the star's given T$_{\rm eff}$ obtained from the PARSEC grid. A more accurate approach to compute this energy would convolve the full stellar spectrum and the frequency-dependent absorption cross section. We tested this approach using the frequency-dependent photoionization cross sections from \citet{1996ApJ...465..487V}.\footnote{Source code containing the analytic fits given in \citet{1996ApJ...465..487V} was obtained from \texttt{http://www.pa.uky.edu/~verner/photo.html}} The blackbody approximation is accurate to within 5\%, yet substantially easier to compute on the fly, as integrals over the blackbody spectrum can be expressed as infinite series that rapidly converge to high precision. Unlike for the ionizing radiation, we assume constant FUV and Lyman-Werner band energies for each star, at 9.8 eV and 12.8 eV respectively.

@@

The left panel of Figure ~\ref{fig:SN histogram} gives the distributions of the peak and average ISM number densities in the SN injection regions for each SN in the simulations presented in this work. As shown, a majority explode in regions at substantially lower densities than the star formation threshold of 200~cm$^{-3}$. For most SNe, $n_{\rm max} \le 1.0$~cm$^{-3}$, which is due to the substantial pre-SN feedback included in our simulations. The right panel of Figure ~\ref{fig:SN histogram} gives the fractional distribution of the calculated radius of the pressure-driven snowplow ($R_{\rm PDS}$) phase of the Sedov-Taylor expansion, adopting the definitions from \citet{Simpson2016}
\begin{equation}
R_{\rm PDS} = 
\begin{cases*}
49.3E^{1/4}_{51}n^{-1/2}_{o} & if  Z $\le$ 0.01 \\
18.5E^{2/7}_{51}n^{-3/7}_{o}Z^{-1/7} & if  Z $\ge$ 0.01 \\
\end{cases*}
\end{equation}
where $E_{51}$ is the injection energy in units of 10$^{51}$~erg, $n_{o}$ is the number density of the medium, and $Z$ is the metallicity in units of $Z_{o}$. \citet{Simpson2016} found that $R_{\rm PDS} > 4.5 \Delta x$ is needed to resolve the SN explosion with thermal energy injection alone, assuming uniform density in the injection region. In practice, the injection region is never uniform. We give $R_{\rm PDS}$ assuming uniform density at both the average and maximum number density in the injection region of each SN. As shown, a majority of SNe are resolved (to the right of the 4.5$\Delta x$ line), but up to 10\% are somewhat under-resolved. In general, these are SN explosions from the most massive (i.e. most prompt) stars. We do not expect that resolving these SNe will dramatically alter our results. However, we note that using our feedback model at resolutions much less than a few pc will require adopting a different injection mechanism.

@@

\begin{figure}
\centering
\includegraphics[width=0.4\linewidth]{sn_density_hist}
\includegraphics[width=0.4\linewidth]{sn_radius_hist}
\caption{{\em Left:} Distribution of peak (orange) and average (black) gas densities in the injection region of each of the SNe from the simulation. {\em Right:} Distribution of the radius of the pressure-driven snowplow phase for each SN assuming a uniform medium at either the peak or average density in the injection region. The vertical lines show one and 4.5 cell widths. This shows that a fraction of these SNe are certainly under resolved, with R$_{\rm PDS}$ less than a single cell size, but the majority are well resolved.}
\label{fig:SN histogram}
\end{figure}

@@

\setcounter{figure}{0}
\section{Cooling and Heating Rates}
\label{appendix:cooling}
\textbf{CHANGE: Section was somewhat confusing in original form. Re-arranged order and re-worded}

@@

For densities where self-shielding is important, we computed \texttt{Cloudy} models of one-dimensional clouds at each temperature and density pair with a physical size corresponding to the Jeans length. In cases where the Jeans length is very large, we limit the size to 100 pc, a reasonable approximation for the maximum size of a self-gravitating cloud. We then adopt the metal line cooling rates obtained from the center of these clouds, whose outside is exposed to the UV background. These models were computed with a modified version of the \texttt{CIAOLoop}\footnote{https://bitbucket.org/brittonsmith/cloudy\_cooling\_tools (our version: https://github.com/aemerick/cloudy\_tools)} code used in \citet{2008MNRAS.385.1443S}. In computing these tables we had to include the cosmic ray ionization that dominates in optically thick clouds. Without this effect, \texttt{Cloudy} becomes unstable in entirely neutral regions. We adopted a cosmic ray ionization rate of 10\% the Milky Way value, though we note that varying this from 1\% to 100\% the Milky Way value did not have an effect on the extracted metal line cooling rates.

Finally, we note that the cooling and heating rates in our simulation depend will deviate from these curves, as they depend upon local conditions that cannot be accounted for here. The primordial cooling and heating rates are computed consistently with the non-equilibrium chemistry solver, leading to deviations from the tabulated primordial rates, and the heating rates depend upon the local stellar radiation field, which are not included in these diagrams.

@@

\textbf{CHANGE: Updated plots with full runs. No significant change to text.}
We demonstrate the effect of varying resolution on three different results in the evolution of our galaxy model in Figure~\ref{fig:resolution_study}, comparing the star formation rate, mass evolution, and metal ejection/retention fractions between high resolution simulations, two runs at 3.6 pc with different initial SN driving rates, and a final run at 7.2 pc. The physics in each of these simulations remains the same with the exception of our star formation threshold. Since star formation is set to occur at gas densities where the Jeans mass is unresolved at the maximum resolution, our star formation algorithm is tied to resolution. A factor of two decrease in resolution translates to a factor of four decrease in the star formation density threshold, from 200~cm$^{-3}$ in our fiducial runs, to 50~cm$^{-3}$ and 12.5~cm$^{-3}$ in our 3.6~pc and 7.2~pc resolution runs respectively.

@@

Figure~\ref{fig:resolution_study} shows that, while decreasing the resolution certainly changes the evolution, it does not do so dramatically. We argue that these differences are within the stochastic variation one would expect from simulations with stochastic star formation and driven mostly by variation in the initial star formation rate of the galaxy. In each of the lower resolution runs, the SFR is larger within the first 100~Myr. This is because, at lower resolution, the initial driving SN are less resolved and less effective at preventing uniform collapse of the gas disk. The 3.6~pc run with a factor of 2 higher initial SN driving rate demonstrates that an increase in the effectiveness of the supernova at lower resolution brings the evolution much closer to that seen at our highest resolution run. We note that the self-consistent SN are still resolved in the 3.6~pc runs since these still explode at the low densities seen in Figure~\ref{fig:SN histogram}. However, although feedback in general is still effective in the 7.2~pc run, SNe are not well resolved at this resolution. This is shown more clearly comparing the left and right panels of Figure~\ref{fig:phase_resolution}, the temperature-density phase diagrams for the 3.6 and 7.2 pc. There is clearly much less hot gas (T $> 10^6$~K) in our lowest resolution run.

The largest qualitative difference between these runs is the H$_2$ mass. As discussed in this work, the H$_2$ mass fraction of our fiducial run is quite high, driven by significant H$_2$ production from gas-phase reactions and the preferential retention of cold, dense H$_2$ gas in feedback driven outflows. However, each of the lower resolution runs produce less H$_2$ (about 40\% less) than our fiducial run. This is certainly an issue in interpreting the generality of our large H$_2$ mass fractions, and requires more tests to confirm. However, it is not unreasonable to expect that the H$_2$ production occurs most effectively at densities not captured in the 3.6~pc and 7.2~pc runs. This suggests that most of our H$_2$ production in the fiducial run occurs in gas with densities above 50~cm$^{-3}$. As shown in Figure~\ref{fig:phase_resolution}, the maximum density is clearly lower in each of the two lower resolution runs,

>>>>

Self-shielding of gas against the metagalactic UV background is important in high-resolution simulations, particularly for low-mass, low-metallicity dwarf galaxies where the UV background is capable of gradually photoevaporating unshielded gas from the galaxy \citep{Simpson2013}. We have implemented the \citet{Rahmati2013} approximate self-shielding method in \texttt{Grackle} to account for \ion{H}{1} self-shielding against the UV background \citep[see][ for more details of this implementation]{GrackleMethod}. We assume \ion{He}{1} ionization generally follows \ion{H}{1}. This allows us to approximate \ion{He}{1} self-shielding using the same form (A. Rahmati, private communication). We ignore \ion{He}{2} photoionization from the UV background entirely. Accounting for self-shielding in this manner leads to an inconsistency in using tabulated, optically-thin metal line cooling rates from \textsc{Cloudy} (see Section 4.1.1 of \citet{Hu2017}). As mentioned previously, we have re-computed metal line cooling tables using \textsc{Cloudy} models of optically thick clouds to be consistent with our self-shielding prescription. This is described in more detail in Appendix ~\ref{appendix:cooling}.

@@

To aid in comparison to other works, we define our ISM phases here, as adopted and modified from \citet{Draine2011}, Table 1.3, and used consistently throughout this analysis. By construction, these phases are mutually exclusive. We take $f_{\rm H_2}$ to be the molecular hydrogen fraction of the total gas mass.

@@

\section{Stellar Radiation Properties}
\label{appendix:radiation}
We 
%mm3 detail 
  determine 
the radiation properties of our model stars as a function of stellar mass and metallicity
%mm3 . The stellar radiation properties are determined 
from the OSTAR2002 grid \citep{Lanz2003} in the regime that it covers, or integrated from an adjusted black body curve for other stars, given the ZAMS stellar radius obtained from the PARSEC \citep{Bressan2012,Tang2014} stellar evolution data set. In Figure~\ref{fig:stellar radiation properties} we plot these properties. We use a constant factor across metallicities for stars not sampled on the OSTAR2002 grid to shift the black body radiation fluxes to be roughly continuous with the ionizing photon rates and luminosities as a function of stellar mass. This requires two factors for each radiation type, one for low mass and one for high mass stars. We use the following multiplicative factors to adjust the black body spectrum for \ion{H}{1} and \ion{He}{1}
%mm3
   ionizing 
radiation respectively: [0.1, 3.2] and [0.0001, 4.0]. We do not find this adjustment to be necessary for the FUV and Lyman-Werner radiation bands.

The ionizing radiation photon energies are taken as the average ionizing photon energy for a black body of the star's given T$_{\rm eff}$ obtained from the PARSEC grid. A more accurate approach to compute this energy would convolve the full stellar spectrum and the frequency-dependent absorption cross section. We tested this approach using the frequency-dependent photoionization cross sections from \citet{1996ApJ...465..487V}.\footnote{Source code containing the analytic fits given in \citet{1996ApJ...465..487V} was obtained from 
%mm3 [changed to \url because \texttt translates the tilde into a space]
\url{http://www.pa.uky.edu/~verner/photo.html}} 
The blackbody approximation is accurate to within 5\%, yet substantially easier to compute on the fly, as integrals over the blackbody spectrum can be expressed as infinite series that rapidly converge to high precision. Unlike for the ionizing radiation, we assume constant FUV and Lyman-Werner band energies for each star, at 9.8 eV and 12.8 eV respectively.

@@

The left panel of Figure~\ref{fig:SN histogram} gives the distributions of the peak and average ISM number densities in the SN injection regions for each SN in our simulations. As shown, a majority explode in regions at substantially lower densities than the star formation threshold of 200~cm$^{-3}$. For most SNe, $n_{\rm max} \le 1.0$~cm$^{-3}$, which is due to the substantial pre-SN feedback included in our simulations. The right panel of Figure ~\ref{fig:SN histogram} gives the fractional distribution of the calculated radius of the pressure-driven snowplow ($R_{\rm PDS}$) phase of the Sedov-Taylor expansion, adopting the definitions from \citet{Simpson2016}
\begin{equation}
R_{\rm PDS} = 
\begin{cases*}
49.3E^{1/4}_{51}n^{-1/2}_{o} & if  Z $\le$ 0.01 \\
18.5E^{2/7}_{51}n^{-3/7}_{o}Z^{-1/7} & if  Z $\ge$ 0.01 \\
\end{cases*}
\end{equation}
where $E_{51}$ is the injection energy in units of 10$^{51}$~erg, $n_{o}$ is the number density of the medium, and $Z$ is the metallicity in units of $Z_{o}$. \citet{Simpson2016} found that $R_{\rm PDS} > 4.5 \Delta x$ is needed to resolve the SN explosion with thermal energy injection alone, assuming uniform density in the injection region. In practice, the injection region is never uniform. We give $R_{\rm PDS}$ assuming uniform density at both the average and maximum number density in the injection region of each SN. As shown, a majority of SNe are resolved (to the right of the 4.5$\Delta x$ line), but up to 10\% are somewhat under-resolved. In general, these are SN explosions from the most massive (i.e. most prompt) stars. We do not expect that resolving these SNe will dramatically alter our results. However, we note that using our feedback model at resolutions much less than a few parsecs will require adopting a different injection mechanism.

@@

\begin{figure}
\centering
\includegraphics[width=0.4\linewidth]{sn_density_hist}
\includegraphics[width=0.4\linewidth]{sn_radius_hist}
\caption{{\em Left:} Distribution of peak (orange) and average (black) gas densities in the injection region of each of the SNe from the simulation. {\em Right:} Distribution of the radius of the pressure-driven snowplow phase for each SN assuming a uniform medium at either the peak or average density in the injection region. The vertical lines show one and 4.5 cell widths. This shows that a fraction of these SNe are certainly underresolved, with R$_{\rm PDS}$ less than a single cell size, 
%mm3
   and somewhat more don't satisfy the 4.5 cell criterion, 
but the majority are well resolved.}
\label{fig:SN histogram}
\end{figure}

@@

\setcounter{figure}{0}
\section{Cooling and Heating Rates}
\label{appendix:cooling}
%mm3 \textbf{CHANGE: Section was somewhat confusing in original form. Re-arranged order and re-worded}

@@

For densities where self-shielding is important, we computed \textsc{Cloudy} models of one-dimensional clouds at each temperature and density pair with a physical size corresponding to the Jeans length. In cases where the Jeans length is very large, we limit the size to 100 pc, a reasonable approximation for the maximum size of a self-gravitating cloud. We then adopt the metal line cooling rates obtained from the center of these clouds, whose outside is exposed to the UV background. These models were computed with a modified version of the \texttt{CIAOLoop}\footnote{https://bitbucket.org/brittonsmith/cloudy\_cooling\_tools (our version: https://github.com/aemerick/cloudy\_tools)} code used in \citet{2008MNRAS.385.1443S}. In computing these tables we had to include the cosmic ray ionization that dominates in optically thick clouds. Without this effect, \textsc{Cloudy} becomes unstable in entirely neutral regions. We adopted a cosmic ray ionization rate of 10\% the Milky Way value, though we note that varying this from 1\% to 100\% the Milky Way value did not have an effect on the extracted metal line cooling rates.

Finally, we note that the cooling and heating rates in our simulation 
%mm3 depend 
will deviate from these curves, as they depend upon local conditions that cannot be accounted for here. The primordial cooling and heating rates are computed consistently with the non-equilibrium chemistry solver, leading to deviations from the tabulated primordial rates, and the heating rates depend upon the local stellar radiation field, which is not included in these diagrams.

@@

%mm3 \textbf{CHANGE: Updated plots with full runs. No significant change to text.}
We demonstrate the effect of varying resolution on three different results in the evolution of our galaxy model in Figure~\ref{fig:resolution_study}, comparing the star formation rate, mass evolution, and metal ejection/retention fractions between our fiducial high-resolution simulation, two runs at 3.6 pc with different initial SN driving rates, and a final run at 7.2 pc. The physics in each of these simulations remains the same with the exception of our star formation threshold. Since star formation is set to occur at gas densities where the Jeans mass is unresolved at the maximum resolution, our star formation algorithm is tied to resolution. A factor of two decrease in resolution translates to a factor of four decrease in the star formation density threshold, from 200~cm$^{-3}$ in our fiducial runs, to 50~cm$^{-3}$ and 12.5~cm$^{-3}$ in our 3.6~pc and 7.2~pc resolution runs respectively.

@@

Figure~\ref{fig:resolution_study} shows that, while decreasing the resolution certainly changes the evolution, it does not do so dramatically. We argue that these differences are within the stochastic variation one would expect from simulations with stochastic star formation and driven mostly by variation in the initial star formation rate of the galaxy. In each of the lower resolution runs, the SFR is larger within the first 100~Myr. This is because, at lower resolution, the initial driving SN are less resolved and less effective at preventing uniform collapse of the gas disk. The 3.6~pc run with a factor of two higher initial SN driving rate demonstrates that an increase in the effectiveness of the supernova at lower resolution brings the evolution much closer to that seen at our highest resolution run. We note that the self-consistent SNe are still resolved in the 3.6~pc runs since these still explode at the low densities seen in Figure~\ref{fig:SN histogram}. However, although feedback in general is still effective in the 7.2~pc run, SNe are not well resolved at this resolution. This is shown more clearly comparing the left and right panels of Figure~\ref{fig:phase_resolution}, the temperature-density phase diagrams for the 3.6 and 7.2 pc. There is clearly much less hot gas (T $> 10^6$~K) in our lowest resolution run.

The largest qualitative difference between these runs is the H$_2$ mass. As discussed in this work, the H$_2$ mass fraction of our fiducial run is quite high, driven by significant H$_2$ production from gas-phase reactions and the preferential retention of cold, dense H$_2$ gas in feedback driven outflows. However, each of the lower resolution runs produce less H$_2$ (about 40\% less) than our fiducial run. This is certainly an issue in interpreting the generality of our large H$_2$ mass fractions, and requires more tests to confirm. However, it is not unreasonable to expect that the H$_2$ production occurs most effectively at densities not captured in the 3.6~pc and 7.2~pc runs. This suggests that most of our H$_2$ production in the fiducial run occurs in gas with densities above 50~cm$^{-3}$. As shown in Figure~\ref{fig:phase_resolution}, the maximum density is clearly lower in each of the two lower resolution runs.
